---
title: "Codebook"
format: html
editor: visual
jupyter: python3
---

## Imports

```{r}
library(tidyverse)
library(readr) 
library(tidymodels)
library(corrplot)
library(discrim)
library(corrplot)
library(Hmisc)
```

## Reading in the Data
```{r}
application_test<-read_csv(here::here("data/application_test.csv")) 
application_train <- read_csv(here::here("data/application_train.csv")) 
bureau_bal <- read_csv(here::here("data/bureau_balance.csv"))
bureau <- read_csv(here::here("data/bureau.csv")) 
ccbal <- read_csv(here::here("data/credit_card_balance.csv")) 
homecredit <- read_csv(here::here("data/HomeCredit_columns_description.csv")) 
cashbal <- read_csv(here::here("data/POS_CASH_balance.csv")) 
prevapp <- read_csv(here::here("data/previous_application.csv")) 
sample_sub <- read_csv(here::here("data/sample_submission.csv"))
```

## Data
### Data Engineering
These are the columns from the main table we decided to keep. Decided to keep 1 score out of 3, drop all that seemed irrelevant or repetitive, and those that had over 15000 nas. The columns kept that did have NA's were external source 2, which had roughly 600 NA's, which we we dropped due to the small number compared to the full data. Also feature engineering of the Region Rating using the two we deemed most important. Moving forward, should we keep the document flags?


```{r}
clean_application_train <- application_train|>
  mutate(REGION_RATING = ((REGION_RATING_CLIENT+REGION_RATING_CLIENT_W_CITY)/2))|>
  select(c(SK_ID_CURR, TARGET, NAME_CONTRACT_TYPE, CODE_GENDER, FLAG_OWN_CAR, FLAG_OWN_REALTY, CNT_CHILDREN, AMT_INCOME_TOTAL, AMT_CREDIT, AMT_ANNUITY, NAME_INCOME_TYPE, NAME_EDUCATION_TYPE, NAME_FAMILY_STATUS, NAME_HOUSING_TYPE, REGION_POPULATION_RELATIVE, DAYS_BIRTH, DAYS_EMPLOYED, DAYS_REGISTRATION, DAYS_ID_PUBLISH, FLAG_PHONE, FLAG_EMAIL, CNT_FAM_MEMBERS, REGION_RATING_CLIENT, REGION_RATING_CLIENT_W_CITY, LIVE_CITY_NOT_WORK_CITY, EXT_SOURCE_2, DAYS_LAST_PHONE_CHANGE, REGION_RATING)) |>
  select(-c(REGION_RATING_CLIENT, REGION_RATING_CLIENT_W_CITY)) |>
  drop_na()

colSums(is.na(clean_application_train))

```

```{r}
app_train <- clean_application_train|>
  select_if(is.numeric)|>
  select(-TARGET)


cor_mat <- cor(app_train)
cor_mat[!lower.tri(cor_mat)]
data.frame(cor_mat) |>
              rownames_to_column() |>
              gather(key="variable", value="correlation", -rowname) |>
              filter(abs(correlation) > 0.5)|>
              filter(abs(correlation) != 1.0)


```

Justin's code for the previous applications table.
```{r}
prev_en <- prevapp |>
  mutate(
    app_to_credit = AMT_APPLICATION / AMT_CREDIT,
    ann_to_credit = AMT_ANNUITY     / AMT_CREDIT,
    valid = NAME_CONTRACT_STATUS == "Approved" &
            AMT_CREDIT  > 0 &
            AMT_ANNUITY > 0 &
            CNT_PAYMENT > 0,
    INTEREST       = ifelse(valid, CNT_PAYMENT * AMT_ANNUITY - AMT_CREDIT, NA_real_),
    INTEREST_RATE  = ifelse(valid, 24 * INTEREST/(AMT_CREDIT * (CNT_PAYMENT + 1)), NA_real_),
    INTEREST_SHARE = ifelse(AMT_CREDIT > 0, INTEREST / AMT_CREDIT, NA_real_)
  )

prev_all <- prev_en |>
  group_by(SK_ID_CURR) |>
  summarise(
    prev_app_count           = n(),
    prev_amt_credit_mean     = mean(AMT_CREDIT, na.rm = TRUE),
    prev_cnt_payment_mean    = mean(CNT_PAYMENT, na.rm = TRUE),
    prev_app_to_credit_mean  = mean(app_to_credit, na.rm = TRUE),
    prev_ann_to_credit_mean  = mean(ann_to_credit, na.rm = TRUE),
    prev_approved_rate       = mean(NAME_CONTRACT_STATUS == "Approved", na.rm = TRUE),
    prev_refused_rate        = mean(NAME_CONTRACT_STATUS == "Refused",  na.rm = TRUE),
    prev_interest_sum        = sum(INTEREST, na.rm = TRUE),
    prev_interest_mean       = mean(INTEREST, na.rm = TRUE),
    prev_rate_mean_all       = mean(INTEREST_RATE, na.rm = TRUE),
    prev_rate_med_all        = median(INTEREST_RATE, na.rm = TRUE),
    prev_share_mean_all      = mean(INTEREST_SHARE, na.rm = TRUE),
    .groups = "drop"
  )


prev_last1 <- prev_en |>
  arrange(SK_ID_CURR, dplyr::desc(.data[["DAYS_DECISION"]])) |>
  group_by(SK_ID_CURR) |>
  slice_head(n = 1) |>
  summarise(
    prev_last1_amt_credit     = first(AMT_CREDIT),
    prev_last1_cnt_payment    = first(CNT_PAYMENT),
    prev_last1_app_to_credit  = first(app_to_credit),
    prev_last1_ann_to_credit  = first(ann_to_credit),
    prev_last1_status         = first(NAME_CONTRACT_STATUS),
    .groups = "drop"
  )

prev_last3 <- prev_en |>
  arrange(SK_ID_CURR, dplyr::desc(.data[["DAYS_DECISION"]])) |>
  group_by(SK_ID_CURR) |>
  slice_head(n = 3) |>
  summarise(
    prev_last3_n                  = n(),
    prev_last3_amt_credit_mean    = mean(AMT_CREDIT,    na.rm = TRUE),
    prev_last3_cnt_payment_mean   = mean(CNT_PAYMENT,   na.rm = TRUE),
    prev_last3_app_to_credit_mean = mean(app_to_credit, na.rm = TRUE),
    prev_last3_ann_to_credit_mean = mean(ann_to_credit, na.rm = TRUE),
    prev_last3_approved_rate      = mean(NAME_CONTRACT_STATUS == "Approved", na.rm = TRUE),
    .groups = "drop"
  )


prev_last5 <- prev_en |>
  arrange(SK_ID_CURR, dplyr::desc(.data[["DAYS_DECISION"]])) |>
  group_by(SK_ID_CURR) |>
  slice_head(n = 5) |>
  summarise(
    prev_last5_n                  = n(),
    prev_last5_amt_credit_mean    = mean(AMT_CREDIT,    na.rm = TRUE),
    prev_last5_cnt_payment_mean   = mean(CNT_PAYMENT,   na.rm = TRUE),
    prev_last5_app_to_credit_mean = mean(app_to_credit, na.rm = TRUE),
    prev_last5_ann_to_credit_mean = mean(ann_to_credit, na.rm = TRUE),
    prev_last5_approved_rate      = mean(NAME_CONTRACT_STATUS == "Approved", na.rm = TRUE),
    .groups = "drop"
  )

prev_first2 <- prev_en |>
  arrange(SK_ID_CURR, .data[["DAYS_DECISION"]]) |>
  group_by(SK_ID_CURR) |>
  slice_head(n = 2) |>
  summarise(
    prev_first2_n                  = n(),
    prev_first2_amt_credit_mean    = mean(AMT_CREDIT,    na.rm = TRUE),
    prev_first2_cnt_payment_mean   = mean(CNT_PAYMENT,   na.rm = TRUE),
    prev_first2_app_to_credit_mean = mean(app_to_credit, na.rm = TRUE),
    prev_first2_ann_to_credit_mean = mean(ann_to_credit, na.rm = TRUE),
    prev_first2_approved_rate      = mean(NAME_CONTRACT_STATUS == "Approved", na.rm = TRUE),
    .groups = "drop"
  )


prev_first4 <- prev_en |>
  arrange(SK_ID_CURR, .data[["DAYS_DECISION"]]) |>
  group_by(SK_ID_CURR) |>
  slice_head(n = 4) |>
  summarise(
    prev_first4_n                  = n(),
    prev_first4_amt_credit_mean    = mean(AMT_CREDIT,    na.rm = TRUE),
    prev_first4_cnt_payment_mean   = mean(CNT_PAYMENT,   na.rm = TRUE),
    prev_first4_app_to_credit_mean = mean(app_to_credit, na.rm = TRUE),
    prev_first4_ann_to_credit_mean = mean(ann_to_credit, na.rm = TRUE),
    prev_first4_approved_rate      = mean(NAME_CONTRACT_STATUS == "Approved", na.rm = TRUE),
    .groups = "drop"
  )


prev_features <- Reduce(function(x, y) left_join(x, y, by = "SK_ID_CURR"),
                        list(prev_all, prev_last1, prev_last3, prev_last5, prev_first2, prev_first4))

application_train_FE <- clean_application_train |> left_join(prev_features, by = "SK_ID_CURR")
```

This still yields a lot of NA's in the final table, so we need to decide how to handle those.
```{r}
colSums(is.na(application_train_FE))
```

Due to aggregation, we now have one row per person. 

Sumanth's code for including bureau information.
Aggregating all of the credit debt information, and creating a debt ratio column. Capping the values at 5 to avoid extreme outliers. 0's if no info for NA's is valid in this case because it means they don't have credit history.
```{r}
clean_bureau <- bureau |>
  select(SK_ID_CURR, AMT_CREDIT_SUM, AMT_CREDIT_SUM_DEBT) |>
  group_by(SK_ID_CURR) |>
  summarise(
    total_credit_sum = sum(AMT_CREDIT_SUM, na.rm = TRUE),
    total_debt_sum = sum(AMT_CREDIT_SUM_DEBT, na.rm = TRUE),
    debt_ratio = ifelse(total_credit_sum > 0, total_debt_sum / total_credit_sum, NA_real_),
    .groups = "drop"
  ) |>
  mutate(debt_ratio = ifelse(is.na(debt_ratio), 0, pmin(debt_ratio, 5)))

colSums(is.na(clean_bureau))
```

```{r}
application_train_FE <- application_train_FE |>
  left_join(clean_bureau |> select(SK_ID_CURR, debt_ratio), by = "SK_ID_CURR") |>
  mutate(debt_ratio = ifelse(is.na(debt_ratio), 0, debt_ratio))

application_train_FE
```


```{r}
colSums(is.na(application_train_FE))
```

### Data Splitting

### Random Split of Data

```{r}
set.seed(123)
application_train_FE$TARGET <- as.factor(application_train_FE$TARGET)
train_size <- floor(0.80 * nrow(application_train_FE))
train_indices <- sample(seq_len(nrow(application_train_FE)), size = train_size)
ran_train_data <- application_train_FE[train_indices, ]
ran_val_data <- application_train_FE[-train_indices, ]
```

### Stratified Split

```{r}
split_strat <- initial_split(application_train_FE, prop = 0.8, strata = "TARGET")
strat_train_data <- training(split_strat)
strat_val_data <- testing(split_strat)
```

### Non Random Split - Represent Test Data

For this split, I am thinking we 

```{r}

```


## Metric Custom Functions

### ROC-AUC, Accuracy, F-1, Specificity, Recall, Precision
```{r}
accuracy <- function(y_hat, y){
    correct <- sum(y_hat == y)
    total <- length(y)
    return(correct / total)
}

precision <- function(y_hat, y){
    tp <- sum((y_hat == 1) & (y == 1))
    fp <- sum((y_hat == 1) & (y == 0))
    precision <- tp / (tp+fp)
    return(precision)
}

recall <- function(y_hat, y){
    tp <- sum ((y_hat == 1) & (y ==1))
    fn <- sum ((y_hat == 0) & (y == 1))
    recall <- tp / (tp+fn)
    return(recall)
}

f1 <- function(y_hat, y){
    precision <- precision(y_hat, y)
    recall <- recall(y_hat, y)
    return (2 * (precision * recall) / (precision + recall))
}

specificity <- function(y_hat, y){
    fp <- sum ((y_hat == 1) & (y == 0))
    tn <- sum ((y_hat == 0) & (y == 0))
    return (tn/(tn+fp))
}

roc_auc <- function(y, y_hat){
    thresholds <- seq(0, 1, by = 0.01)
    tpr <- c()
    fpr <- c()
    for (t in thresholds){
        tp <- sum((y_hat >= t) & (y == 1))
        fn <- sum((y_hat < t) & (y == 1))
        fp <- sum((y_hat >= t) & (y == 0))
        tn <- sum((y_hat < t) & (y == 0))
        tpr <- c(tpr, tp / (tp + fn))
        fpr <- c(fpr, fp / (fp + tn))
    }
    auc <- 0
    for (i in 2:length(tpr)){
        auc <- auc + (fpr[i] - fpr[i-1]) * (tpr[i] + tpr[i-1]) / 2
    }
    return(auc)
}

```

### Cross Validation Function
```{r}
cross_val <- function(model, formula, data, k = 5, costVal=NULL, marginVal=NULL){
    set.seed(23)
    folds <- sample((rep(1:k, length.out = nrow(data))))
    metrics <- data.frame(Accuracy = numeric(k+1),
                          F1 = numeric(k+1),
                          Precision = numeric(k+1),
                          Recall = numeric(k+1),
                          Specificity = numeric(k+1),
                          ROC_AUC = numeric(k+1))
    for (i in 1:k){
        train_fold = data[folds != i, ]
        val_fold = data[folds == i,]
        if (deparse(substitute(model)) == "SVC_Model") {
          y_hat <- model(trainData = train_fold, valData = val_fold, formula = formula, costVal=0.1, marginVal=1)
        } else {
          y_hat <- model(trainData = train_fold , valData = val_fold, formula = formula)
        }
        y_hat <- as.numeric(as.character(y_hat$.pred_class))
        y <- as.numeric(as.character(val_fold$TARGET))
        metrics$Accuracy[i] <- accuracy(y_hat, y)
        metrics$Precision[i] <- precision(y_hat, y)
        metrics$Recall[i] <- recall(y_hat, y)
        metrics$F1[i] <- f1(y_hat, y)
        metrics$Specificity[i] <- specificity(y_hat, y)
        metrics$ROC_AUC[i] <- roc_auc(y, y_hat)
    }
    metrics$Accuracy[k+1] <- mean(metrics$Accuracy)
    metrics$Precision[k+1] <- mean(metrics$Precision)
    metrics$Recall[k+1] <- mean(metrics$Recall)
    metrics$F1[k+1] <- mean(metrics$F1)
    metrics$Specificity[k+1] <- mean(metrics$Specificity)
    metrics$ROC_AUC[k+1] <- mean(metrics$ROC_AUC)
    return(metrics)
}
```


## Models
### Logistic Regression
```{r}
library(MASS)
generateModelFormula <- function(yResponse,xPredictors){
  formula <- as.formula(paste(yResponse,"~",paste(xPredictors,collapse="+")))
  return(formula)
}
```

```{r}
yResponse="TARGET"
xPredictors=c("AMT_INCOME_TOTAL","AMT_CREDIT","AMT_ANNUITY","AMT_GOODS_PRICE")
formula <- generateModelFormula(yResponse,xPredictors)
```


```{r}
# logistic regression model function
logistic_regression_model <- function(trainData, valData, formula){
  fitted_logistic_model <- logistic_reg() %>%  
    set_engine("glm") %>% 
    set_mode("classification")%>%
    fit(formula,data=trainData)
  pred_class <- predict(fitted_logistic_model, new_data=valData, type="class")
  return(pred_class)
}
```

```{r}
# run logistic regression
yResponse="TARGET"
xPredictors=c("AMT_INCOME_TOTAL","AMT_CREDIT","AMT_ANNUITY")
formula <- generateModelFormula(yResponse,xPredictors)

strat_model_lr <- cross_val(logistic_regression_model, formula, strat_train_data)
strat_model_lr
ran_model_lr <- cross_val(logistic_regression_model, formula, ran_train_data)
```

### Support Vector Machine Classification Model

```{r}
#SVC Model function
SVC_Model <- function(trainData, valData, formula, costVal, marginVal){
  svm_model <- svm_linear( cost=costVal, margin=marginVal) %>%
    set_mode("classification") %>% 
    set_engine("LiblineaR")
  svmModel_fit <- svm_model %>% fit(formula, data=trainData)
  return(predict(svmModel_fit, valData, type="class"))
}
```

```{r}
#run SVMClassification
costVal=0.1
marginVal=1
strat_model_svc <- cross_val(SVC_Model, formula, strat_train_data, costVal=costVal, marginVal=marginVal)
strat_model_svc
ran_model_svc <- cross_val(SVC_Model, formula, ran_train_data, costVal=costVal, marginVal=marginVal)
```


### LDA: Linear Discriminant Analysis Model
```{r}
#LDA: Linear Discriminant Analysis Model
LDA_model <- function(trainData, valData, formula){
  discrim_model <- discrim_linear(
    mode="classification",
    penalty=NULL,
    regularization_method=NULL,
    engine="MASS"
  )
  discrim_modelFit <- discrim_model %>% fit(formula, data=trainData)
  return(predict(discrim_modelFit, valData, type="class"))
}
```

```{r}
#run LDA
strat_model_lda <- cross_val(LDA_model, formula, strat_train_data)
strat_model_lda
ran_model_lda <- cross_val(LDA_model, formula, ran_train_data)
```
